---
title             : "papaja: Table examples"
shorttitle        : "Table examples"

author: 
  - name          : "Frederik Aust"
    affiliation   : ""
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"

affiliation:
  - id            : ""
    institution   : "University of Cologne"

abstract: |
  This document showcases some `apa_table()` examples.
  
bibliography      : "apa_table_example.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
library("psych")
r_refs("r-references.bib")

knitr::opts_chunk$set(echo = TRUE)
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r load-data}
load("../../src/slides/cosmetic_surgery.Rdata")
```

```{r fit-models}
lm1 <- lm(formula = Post_QoL ~ BDI, data = cosmetic_surgery)
lm2 <- lm(formula = Post_QoL ~ Base_QoL + BDI, data = cosmetic_surgery)
lm3 <- lm(formula = Post_QoL ~ Base_QoL + Age + Gender + BDI, data = cosmetic_surgery)

lm3_apa <- apa_print(lm3)

mod_comp_apa <- apa_print(
  list(Bivariate = lm1, Baseline = lm2, Demographics = lm3)
  , anova_fun = anova
  , boot_samples = 0
)
```

(ref:regression-table-caption) Results of model comparisons between three set of linear regression models that describe participants' quality of life post surgery.

(ref:regression-table-note) The predictor *Gender* was dummy coded with females as reference category. BDI = Beck's depression inventory [@bdi], *Bas QoL* = Quality of life at baseline.

```{r regression-table-syntax}
apa_table(
  lm3_apa$table
  , caption = "(ref:regression-table-caption)"
  , note = "(ref:regression-table-note)"
)
```

(ref:mod-comp-table-caption) Results of model comparisons between three set of linear regression models that describe participants' quality of life post surgery.

(ref:mod-comp-table-note) The predictor *Gender* was dummy coded with females as reference category. BDI = Beck's depression inventory [@bdi], *Bas QoL* = Quality of life at baseline, AIC = Akaike information criterion, BIC = Bayesian information criterion.

```{r standard-table, results = "asis"}
apa_table(
  mod_comp_apa$table
  , align = "lrrr"
  , caption = "(ref:mod-comp-table-caption)"
  , note = "(ref:mod-comp-table-note)"
  , escape = TRUE
)
```

```{r more-elaborate-table, results = "asis"}
apa_table(
  mod_comp_apa$table
  , align = "lrrr"
  , caption = "(ref:mod-comp-table-caption)"
  , note = "(ref:mod-comp-table-note)"
  , escape = FALSE
  , col_spanners = list("Covariate models" = c(3, 4))
  , stub_indents = list("Predictors" = 1:5, "Model fit" = 6:12, "Model comparison" = 13:19)
  , midrules = c(6, 14)
  # , font_size = "tiny"
  , longtable = TRUE
)
```

```{r descriptives}
descriptives <- describeBy(
  subset(cosmetic_surgery, select = -c(ID, Clinic, Surgery, Gender, Reason))
  , cosmetic_surgery$Reason
)

class(descriptives)

descriptives <- unclass(descriptives)
is.list(descriptives)

# Add variable labels to each data.frame in the list
descriptive_labels <- c(
  # vars = "Variable"
  n = "$n$"
  , mean = "$\\overline{M}$"
  , sd = "$SD$"
  , median = "$\\widetilde{M}$"
  , trimmed = "$\\overline{M_t}$"
  , mad = "MAD"
  , min = "Min"
  , max = "Max"
  , range = "Range"
  , skew = "$\\gamma_1$"
  , kurtosis = "$\\gamma_4$"
  , se = "$SE$"
)

descriptives[[1]]$n <- as.integer(descriptives[[1]]$n)
descriptives[[1]]$vars <- NULL
descriptives[[2]]$n <- as.integer(descriptives[[2]]$n)
descriptives[[2]]$vars <- NULL

variable_labels(descriptives[[1]]) <- descriptive_labels
variable_labels(descriptives[[2]]) <- descriptive_labels
```

(ref:descriptives-table) Predictor descriptives seperately for reported reason for surgery.

```{r merged-table, results = "asis"}
apa_table(
  descriptives
  , align = c("l", rep("r", 12))
  , caption = "(ref:descriptives-table)"
  , added_stub_head = "Variable"
  , midrules = c(1, 5, 6)
  , landscape = TRUE
)
```


\newpage

# References

::: {#refs}
:::
