---
title: "Exploring Gapminder data"
author: "R User"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)

options(scipen = 10,
        digits = 2)

library(tidyverse)
library(gapminder)
```

# The data

In this report, we explore data from [*Gapminder*](https://www.gapminder.org/). More specifically, we use an excerpt from these data that is included in the [`gapminder` package for `R`](https://github.com/jennybc/gapminder). The data contains information about life expectancy, population, and GDP per capita for `r n_distinct(gapminder$country)` countries from `r n_distinct(gapminder$continent)` different continents, spanning the time from `r min(gapminder$year)` to `r max(gapminder$year)`.

# Analyses

For our exploration, we will focus on data from the year 2007.

```{r wrangle-data, echo=FALSE}
gap_2007 <- gapminder %>% 
  filter(year == 2007)
```

## Life expectancy

In 2007, life expectancy ranged from `r min(gapminder$lifeExp)` to `r max(gapminder$lifeExp)`, with a global average of `r mean(gapminder$lifeExp)`.

```{r lifeExp-dist, fig.cap="Distribution of life expectancy in 2007"}
gap_2007 %>% 
  ggplot(aes(x = lifeExp)) +
  geom_density(fill="#69b3a2",
               color="#e9ecef") +
  scale_y_continuous(expand=expansion(mult=c(0,0.1))) +
  scale_x_continuous(expand=expansion(mult=c(0,0))) +
  theme_minimal()
```

```{r gap-toplife}
gap_2007 %>%
  select(country, continent,lifeExp) %>%
  arrange(-lifeExp) %>% 
  head(10) %>% 
  knitr::kable(caption = "10 countries with the highest life expectancy in 2007")
```

```{r box-life, fig.cap="Differences between continents in average life expectancy in 2007"}
gap_2007 %>% 
  ggplot(aes(x = continent, y = lifeExp)) +
  geom_boxplot(outlier.colour = "hotpink") +
  geom_jitter(position = position_jitter(width = 0.1, height = 0),
              alpha = 0.25)
```

## GDP per capita

The minimum GDP per capita in 2007 was `r min(gapminder$gdpPercap)`, the maximum was `r max(gapminder$gdpPercap)`, and the worldwide mean was `r mean(gapminder$gdpPercap)`.

As we can see, GDP per capita is heavily right-skewed.

```{r gdp-dist, fig.cap="Distribution of GPD per capita in 2007"}
gap_2007 %>% 
  ggplot(aes(x = gdpPercap)) +
  geom_density(fill="#69b3a2",
               color="#e9ecef") +
  scale_y_continuous(expand=expansion(mult=c(0,0.1))) +
  scale_x_continuous(expand=expansion(mult=c(0,0))) +
  theme_minimal()
```

```{r gap-topgdp}
gap_2007 %>%
  select(country, continent,gdpPercap) %>%
  arrange(-gdpPercap) %>% 
  head(10) %>% 
  knitr::kable(caption = "10 countries with the highest GDP per capita in 2007")
```

```{r box-gdp, fig.cap="Differences between continents in GDP per capita in 2007"}
gap_2007 %>% 
  ggplot(aes(x = continent, y = gdpPercap)) +
  geom_boxplot(outlier.colour = "hotpink") +
  geom_jitter(position = position_jitter(width = 0.1, height = 0),
              alpha = 0.25)
```

## Relationship between life expectancy and GDP per capita

We employ a bubble plot as made popular by the [TED talk by Hans Rosling](https://www.ted.com/talks/hans_rosling_the_best_stats_you_ve_ever_seen) to visualize the relationship between GDP per capita and life expectancy in 2007. The plot also includes visual information about the continent on which a country is located as well as its population size.

```{r bubble}
gap_2007 %>% 
ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(size = pop,
                 color = continent),
             alpha = .5) +
  scale_x_log10() +
  labs(caption = "Due to the heavily skewed distribution of GDP per capita, the X-axis uses a log-scale")
```

# Reproducibility information

This document was created with the following settings:

```{r session-info, echo=FALSE}
sessionInfo() %>% 
  print(locale = FALSE)
```
